---
title: "In Class Exercise 1"
author: "Joshua TING"
date: "August 26, 2024"
date-modified: "last-modified"
format:
  html:
    code-fold: true
    code-summary: "Code Chunk"
    number-sections: true
execute:
  eval: true #r will run through all codes
  echo: true #r will display all code chunk
  warning: false #for mark down
  freeze: true #r will not render all existing  html files
  message: false #avoid printing warning message
editor: source
---

# **In-Class Exercise 1**

## **Installing Packages**

```{r}
pacman::p_load(sf, tidyverse, tmap, ggstatsplot)
```

## **Reading Data**

### Importing shapefile

```{r}
mpsz14_shp = st_read(dsn = "data/", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

### Importing kml file

```{r}
#| eval: false
mpsz14_kml <- st_read("data/MasterPlan2014SubzoneBoundaryWebKML.kml")
```

### Troubleshooting: Resolving the error of corrupted kml file

```{r}
st_write(mpsz14_shp,
         "data/MP14_SUBZONE_WEB_PL.kml",
         delete_dsn = TRUE) #delete the existing file with same name first
```

### To import shapefile Master Plan 2025

```{r}
mpsz19_shp = st_read(dsn = "data/", 
                  layer = "MPSZ-2019")
```

### To import kml Master Plan 2025

```{r}
mpsz19_shp <-  st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

::: callout-note
## PCS over GCS

To use Projected Coordinate System (PCS) as geography is flattened and measured using (i.e. metres) instead of Geography Coordinate System (GCS) as it is not accurate. For an example, one degree in the equator and north pole makes a significant difference and GCS will distort it.
:::

## Transforming coordinate system from GCS to PCS

This mathematically converts one coordinate to another coordinate.

```{r}
mpsz19_shp <- st_read(dsn = "data/",
                      layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

::: callout-note
## Choosing CSV over XLS

CSV. over XLS as there are multiple hindrances (i.e. extra space, symbols) that requires extra effort to clean. Use text files without structures.
:::

## New Practice

### Reading the CSV data

```{r}
popdata <- read.csv("data/respopagesextod2023/respopagesextod2023.csv")
```

### Aggregation the data

```{r}
#| eval: false
popdata2023 <- popdata %>%
  group_by(PA, SZ, AG) %>% #aggregation
  summarise('POP'=sum('Pop')) %>%
  ungroup() %>%
  pivot_wider(names_from=AG,
              values_from=POP)
```

```{r}
#| eval: false
popdata2023 <- popdata %>%
  mutate(YOUNG=rowSums(.[3:6]))
```
