---
title: "In-Class Exercise 2"
author: "Joshua TING"
date: "September 1, 2024"
date-modified: "last-modified"
format:
  html:
    code-fold: true
    code-summary: "Code Chunk"
    number-sections: true
execute:
  eval: false #r will run through all codes
  echo: true #r will display all code chunk
  warning: false #for mark down
  freeze: true #r will not render all existing  html files
  message: false #avoid printing warning message
editor: source
---

In Class Exercise 2

Maptools is retired and binary is removed from CRAN. If you would want to use it, you can retrieve from CRAN. However, R version must support the package.

```{r}
install.packages("maptools",
                 repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

```{r}
pacman::p_load()
```

Issue 2: Creating coastal outline

in sf package, there are two functions allow us to combine multiple simple features into one simple features. They are st_combine and st_union().

sf: <https://r-spatial.github.io/sf/>

Working with st_union()

The code chunk below st_union() is used to derive the coastal outline sf tibble data.frame.

```{r}
sg_sf <- mpsz_sf %>%
  st_union()
```

sg_sf will look similar to the figure below.

```{r}
print(sg_sf)
```

as.ppp() allows to create point object and as.owin creates boundary.

\`\`\`{r}\

```{r}

par(bg="#E4D5C9")

gridded_kde_childcareSG_ad <- as(
  kde_childcareSG_adaptive,
  "SpatialGridDataFrame")
spplot(gridded_kde_childcareSG_ad)
```

::: {.callout-tip appearance="minimal"}
## Monte Carlo Simulation

In order to ensure reproducibility, it's important to include the code chunk below before using spatstat functions involve Monte Carlo Simulation.
:::

```{r}
set.seed(1234)

```

1.  ensure km are m in PCS
2.  
